Spectral Lags in Very High Energy Emission of Gamma Ray Bursts

### 2013-11-21 gammaRays.tex
Small style improvements

To understand the observed spectral lag, lets explore geometry of the jet. First of all, lets make some assumptions:
\begin{enumerate}
\item{Time $t = 0$, a spherical shell of plasma is emitted. The center is called the central engine.}
\item{The shell points propagate with a constant velocity $v \sim 1$, so at the time $t$ the radius of the shall is $v t$.}
\item{Each point of the shell is an isotropic radiator in its reference frame.}
\item{
	The radiation intensity is a function of the radiator position and the radiation frequency:
	\begin{equation}
		\eta\left(r,\theta,\omega\right) = 
		\frac{\eta_0}{1 + \left(\frac{r}{r_0}\right)^n}
		\exp\left(
			-\left(\frac{\theta}{\theta_0}\right)^2
			\left(\frac{\omega}{\omega_0}\right)^{-2k}
		\right)
		\left(\frac{\omega}{\omega_0}\right)^\alpha
	\end{equation}
	$\eta$ is a number of particles emitted per volume per solid angle per frequency. It is a function of the distance $r$ from the central engine, of the off-axis angle $\theta$, and of the radiation frequency $\omega$.
}
\end{enumerate}

@@

Lets begin by calculating a time at which some particular photon is observed. This time is a function of the radiator location $\left(r, \theta, \phi\right)$, as well as the observer location $\left(d, \chi, 0\right)$. (Here we choose coordinates so that the rotation angle of the observer is $0$.) Lets assume for now that the observer is far enough to resolve the jet geometry, yet close enough so that the expansion of space is negligible. This assumption allows us to ignore the effects of cosmology for this calculation.

@@

We also need to know the areas of two spheres: the photon sphere, the sphere over which the photons emitted in a particular burst are spread; and the bursts sphere, the sphere over which the bursts at a particular redshift are distributed.

@@

The remaining integrals over $\theta$ and $\chi$ are hard to do symbolically, so we compute them numerically. To optimize this computation we can use the assumption of small $\theta$ and $\chi$, so that:
\begin{align*}
\sin\theta &\sim \theta \\
\cos\sigma\left(\theta, \phi, \chi \right) &= \cos\theta \cos\chi + \sin\theta \sin\chi \cos\phi \sim 1 - \frac{\theta^2}{2} - \frac{\chi^2}{2} + \theta\chi\cos\phi
\end{align*}
Using that, and the fact that this integral is an even function of $\theta$, we arrive to the optimized expressions for $p$, $I$ and $J$:
\begin{align}
p\left( z, \chi; \tau_1, \tau_2; \omega_1, \omega_2 \right) &= \frac{A_\text{det}}{A_\text{ph}\left(z\right)}
\frac{2\eta_0}{\left(v\gamma\left(1+z\right)\right)^{2-\alpha}}
\int_0^{\infty} \dif \theta \int_0^{\pi} \dif \phi \frac{\theta}{\left(\frac{1}{v} - 1 + \frac{\theta^2}{2} + \frac{\chi^2}{2} - \theta\chi\cos\phi\right)^{5-\alpha}}\nonumber\\
&\qquad {} \times \left( I\left(z,\chi,\omega_2; \theta,\phi\right) - I\left(z,\chi,\omega_1; \theta,\phi\right) \right) \left( J\left(z,\chi,\tau_2; \theta,\phi\right) - J\left( z,\chi,\tau_1; \theta,\phi \right) \right) \\
I\left(z,\chi,\omega;\theta,\phi\right) &= \frac{
	\omega
	\left(\frac{\omega}{\omega_0}\right)^{\alpha }
	E_{\frac{\alpha +1}{2 k}+1}\left(
		\left(\frac{\theta}{\theta_0}\right)^2
		\left(\frac{\omega}{\omega_0}\right)^{-2 k}
		\left(
			v \gamma \left(1+z\right) \left(\frac{1}{v} - 1 + \frac{\theta^2}{2} + \frac{\chi^2}{2} - \theta\chi\cos\phi \right)
		\right)^{-2 k}
	\right)
}{
	2 k
} \\
J\left(z,\chi,\tau;\theta,\phi\right) &= \frac{\tau^3}{3} \,
_2F_1\left(
	1,
	\frac{3}{n};
	\frac{n+3}{n};
	-\left(\frac{\tau}{
		r_0 \left( \frac{1}{v} -  1 + \frac{\theta^2}{2} + \frac{\chi^2}{2} - \theta\chi\cos\phi \right) \left(1+z\right)
	}\right)^n
\right)
\end{align}

@@

\subsection{Spectrum}
\subsection{Distribution}
\subsection{Case study}

>>>>

To understand the observed spectral lag, lets explore geometry of the jet. First of all, lets make some assumptions:
\begin{enumerate}
\item{Time $t = 0$, a spherical shell of plasma is emitted. The center is called the central engine.}
\item{The shell points propagate with a constant velocity $v \sim 1$, so at the time $t$ the radius of the shall is $v t$.}
\item{Each point of the shell is an isotropic radiator in its rest frame.}
\item{
	The radiation intensity is a function of the radiator position and the radiation frequency:
	\begin{equation}
		\eta\left(r,\theta,\omega\right) = 
		\frac{\eta_0}{1 + \left(\frac{r}{r_0}\right)^n}
		\exp\left(
			-\left(\frac{\theta}{\theta_0}\right)^2
			\left(\frac{\omega}{\omega_0}\right)^{-2k}
		\right)
		\left(\frac{\omega}{\omega_0}\right)^\alpha
	\end{equation}
	$\eta$ is a number of particles emitted per volume per solid angle per frequency. It is a function of the distance $r$ from the central engine, of the off-axis angle $\theta$, and of the radiation frequency $\omega$.
}
\end{enumerate}

@@

Lets begin by computing a time at which some particular photon is observed. This time is a function of the radiator location $\left(r, \theta, \phi\right)$, as well as the observer location $\left(d, \chi, 0\right)$. (Here we choose coordinates so that the rotation angle of the observer is $0$.) Lets assume for now that the observer is far enough to resolve the jet geometry, yet close enough so that the expansion of space is negligible. This assumption allows us to ignore the effects of cosmology for this calculation.

@@

We also need to know the areas of two spheres: the photon sphere, a sphere over which photons emitted in a particular burst are spread; and the bursts sphere, a sphere over which bursts at a particular redshift are distributed.

@@

The remaining integrals over $\theta$ and $\chi$ are hard to do symbolically, so we compute them numerically. To optimize this computation we can use the assumption of small $\theta$ and $\chi$, so that:
\begin{align*}
\sin\theta &\sim \theta \\
\cos\sigma\left(\theta, \phi, \chi \right) &= \cos\theta \cos\chi + \sin\theta \sin\chi \cos\phi \sim 1 - \frac{\theta^2}{2} - \frac{\chi^2}{2} + \theta\chi\cos\phi
\end{align*}
Using that, and the observation that this integral is an even function of $\theta$, we arrive to the optimized expressions for $p$, $I$ and $J$:
\begin{align}
p\left( z, \chi; \tau_1, \tau_2; \omega_1, \omega_2 \right) &= \frac{A_\text{det}}{A_\text{ph}\left(z\right)}
\frac{2\eta_0}{\left(v\gamma\left(1+z\right)\right)^{2-\alpha}}
\int_0^{\infty} \dif \theta \int_0^{\pi} \dif \phi \frac{\theta}{\left(\frac{1}{v} - 1 + \frac{\theta^2}{2} + \frac{\chi^2}{2} - \theta\chi\cos\phi\right)^{5-\alpha}}\nonumber\\
&\qquad {} \times \left( I\left(z,\chi,\omega_2; \theta,\phi\right) - I\left(z,\chi,\omega_1; \theta,\phi\right) \right) \left( J\left(z,\chi,\tau_2; \theta,\phi\right) - J\left( z,\chi,\tau_1; \theta,\phi \right) \right) \\
I\left(z,\chi,\omega;\theta,\phi\right) &= \frac{
	\omega
	\left(\frac{\omega}{\omega_0}\right)^{\alpha }
	E_{\frac{\alpha +1}{2 k}+1}\left(
		\left(\frac{\theta}{\theta_0}\right)^2
		\left(\frac{\omega}{\omega_0}\right)^{-2 k}
		\left(
			v \gamma \left(1+z\right) \left(\frac{1}{v} - 1 + \frac{\theta^2}{2} + \frac{\chi^2}{2} - \theta\chi\cos\phi \right)
		\right)^{-2 k}
	\right)
}{
	2 k
} \\
J\left(z,\chi,\tau;\theta,\phi\right) &= \frac{\tau^3}{3} \,
_2F_1\left(
	1,
	\frac{3}{n};
	\frac{n+3}{n};
	-\left(\frac{\tau}{
		r_0 \left( \frac{1}{v} -  1 + \frac{\theta^2}{2} + \frac{\chi^2}{2} - \theta\chi\cos\phi \right) \left(1+z\right)
	}\right)^n
\right)
\end{align}

@@

\subsection{Distribution}
\subsection{Burst fraction with observable high-energy photons}
